From noise to knowledge: how randomness generates novel phenomena and reveals information

### 2018-03-06 paper/paper.Rmd
fix trailing whitespace

We turn from stochastic resonance to oscillations of a  different
sort, illustrated in Figure \ref{noisy_switch} in empirical data on 
rubella outbreaks in Copenhagen, analyzed by @Keeling2001, and in 
simulations of a standard model shown in Figure \ref{noisy_switch}B.
This pattern is characterized by long stretches during
which fluctuations in the state variable remain concentrated around
a high average value followed by sudden, rapid transitions into
fluctuations concentrated around a lower value.  This behavior
arises through the interaction of stochasticity and alternative
stable state dynamics.

Both for clarity of presentation and highlight the generality of 
stochastic switching in diverse models, we will focus not specifically
on the more complex disease model analyzed in @Keeling2001 but
rather rely on the classic resource consumption model of @May1977
which has been widely used to study alternative state dynamics.
As we shall see, stochastic switching is a generic property of 
models with alternative stable states which will have other implications
later on.  In @May1977 model, a resource species grows according to a logistic growth
model and is consumed according to a Holling Type-III functional
response curve, to which I have simply introduced the addition of
environmental noise:

>>>>

We turn from stochastic resonance to oscillations of a  different
sort, illustrated in Figure \ref{noisy_switch} in empirical data on 
rubella outbreaks in Copenhagen, analyzed by @Keeling2001, and in 
simulations of a standard model shown in Figure \ref{noisy_switch}B.
This pattern is characterized by long stretches during
which fluctuations in the state variable remain concentrated around
a high average value followed by sudden, rapid transitions into
fluctuations concentrated around a lower value.  This behavior
arises through the interaction of stochasticity and alternative
stable state dynamics.  To sidestep some of the complexity of the disease
model proposed in @Keeling2001 we will illustrate the same stochastic 
switching phenomenon in the classic resource consumption model of @May1977
which has been widely used to study alternative state dynamics.
As we shall see, stochastic switching is a generic property of 
models with alternative stable states which will have other implications
later on.  In @May1977 model, a resource species grows according to a logistic growth
model and is consumed according to a Holling Type-III functional
response curve, to which I have simply introduced the addition of
environmental noise:
